evidence of adaptation-independent time-warped spike timing noise correlations in cortical networks

### 2020-08-10 AnalysisOutputLatex.tex
Lunch!

%<*ClusterCorrelationSummary1>
Flat clusters were linearly correlated above chance levels (Fisher's method: $\chi^{2}(7844, N = 3922) = 15380.2, $ $p < .001$; Fig. 3e). Of the final angled clusters, 97.6\% (160/164) were positively correlated (Fig. 3f), suggesting the presence of a factor affecting response latency monotonically across the population. For the corresponding flat cluster of the criteria fulfilling angled clusters, spike pair differences were correlated with the spike time of the first unit in the pair (Fisher's method: $\chi^{2}(328, N = 164) = 8366.9, $ $p < .001$; Fig. 3g), showing that correlations were not explained by a model of the form $s_1 = s_0 + d + independent\_noise$ where $d$ is a fixed difference.
%</ClusterCorrelationSummary1>

%<*ClusterCorrelationSummary2>
For the final angled clusters, 81.7\% (134/164) of the corresponding flat clusters had linear regression p-values less than the p-values of linear regression applied to the unclustered first spike pairs (Fig. 3h). This implies that without clustering, linear regression would miss a large proportion of first spike correlations.
%</ClusterCorrelationSummary2>

%<*AngledClustersUniquePairsSummary>
Within this time window, the algorithm extracted a total of 164 angled clusters from 164 unique response distributions, such that there were no repeated or similar clusters.
%</AngledClustersUniquePairsSummary>

%<*OriginalAnglesSummary>
For the positively correlated angled clusters, Fig. 4a shows the distribution of the mean $\theta_{45}$ angles with 95\% confidence intervals. 66.5\% (109/164) of angles were significantly greater than $0^{\circ}$ (p-value < 0.025) and less than $45^{\circ}$ (p-value < 0.025). 
%</OriginalAnglesSummary>

%<*OriginalTestsPassedSummary>
Of the 164 angled clusters, 44 were determined to be stationary and sufficiently correlated by fulfilling the criteria described in the Methods \ref{methods:angle_and_factor_analysis_criteria}, such that an underlying factor could underlie the covariation.
%</OriginalTestsPassedSummary>

%<*OriginalTestsPassedAngleSummary>
Of the 44 criteria fulfilling angled clusters, 26 had estimated angles significantly different from both $0^{\circ}$ (p-value < 0.025; Fig. 4f) and $45^{\circ}$ (p-value < 0.025; Fig. 4f). Of the 44 criteria fulfilling angled clusters, normality was not rejected for 18 clusters (Henze-Zirkler p > 0.05), of which 11 had an estimated angle significantly different from both $0^{\circ}$ (p-value < 0.025) and $45^{\circ}$ (p-value < 0.025).
%</OriginalTestsPassedAngleSummary>

%<*OriginalTestsPassedFactorAnalysisSummary>
Factor analysis could be applied to the 44 criteria fulfilling angled clusters, to give an estimate of the factor-conditioned cluster response distributions for the stationary angled clusters.
%</OriginalTestsPassedFactorAnalysisSummary>

%<*ClusterSingleUnitKPSSStationaritySummary>
For the 164 angled clusters, 118 of the corresponding 328 ($=2*164$) single unit cluster first spike sequences were determined as KPSS non-stationary (p < 0.05; Fig. 4d).
%</ClusterSingleUnitKPSSStationaritySummary>

%<*ClusterSingleUnitTrialIndexLRSummary>
The cluster spike times of either 0, 1 or 2 of the units in a cluster pair were correlated with trial index (Fisher's method: $\chi^{2}(656, N = 328) = 2551.6, $ $p < .001$; Fig. 4b,c).
%</ClusterSingleUnitTrialIndexLRSummary>

%<*SelectivelyDifferencedSummary>
In total, 87 clusters had differencing applied to atleast one unit. 87 of these clusters fulfilled the criteria, giving a total of 87 criteria fulfilling angled clusters (including the undifferenced stationary clusters). 
%</SelectivelyDifferencedSummary>

%<*SelectivelyDifferencedCorrelatedSummary>
In total, 59 of the 87 criteria fulfilling clusters were linearly correlated following differencing (p-value < 0.05) (Fig. 6h). 26 of the correlated criteria fulfilling clusters had estimated angles significantly greater than $0^{\circ}$ (p-value < 0.025) and less than $45^{\circ}$ (p-value < 0.025). Of these, normality was not rejected for 15 clusters (Henze-Zirkler p > 0.05). Fig. 6j shows that the distribution of estimated angles for the successfully differenced clusters. Fig. 8? shows the difference between original and differenced estimated clusters?
%</SelectivelyDifferencedCorrelatedSummary>

%<*SelectivelyDifferencedTestsPassedNewSummary>
In total, 87 clusters had differencing applied to atleast one unit. 41 of these were correlated and fulfilled the criteria. 26 of these clusters had estimated angles significantly greater than $0^{\circ}$ (p-value < 0.025; Fig 6b) and less than $45^{\circ}$ (p-value < 0.025; Fig 6b), of which 14 normality was not rejected for (Henze-Zirkler p < 0.05). These angles are similar to the original angles before differencing (Fig. 6c), showing that the differencing operation does not largely alter cluster angle. Fig. 6d shows the amount of variance explained by the pairwise correlation after trend removal. 
%</SelectivelyDifferencedTestsPassedNewSummary>

%<*SelectivelyDifferencedTestsPassedEigenvalueSummary>
The factor analysis model from above can be applied to the 41 correlated criteria fulfilling differenced clusters, as the factor correlation matrices of each cluster only had one eigenvalue greater than 1. 
%</SelectivelyDifferencedTestsPassedEigenvalueSummary>

%<*SelectivelyDifferencedBoxJenkinsDifferencedSummary>
85 of the 87 differenced clusters had an ARMA model applied to atleast one unit. Almost all ARMA models fit for differenced clusters were of the MA type (Fig 7a), as expected for the negative autocorrelations introduced by differencing. 
%</SelectivelyDifferencedBoxJenkinsDifferencedSummary>

%<*SelectivelyDifferencedBoxJenkinsCorrelatedSummary>
After the described application of the Box-Jenkins method, there were 44 criteria fulfilling clusters. Normality was not rejected for 18 of these clusters (Henze-Zirkler p > 0.05). 26 of the correlated criteria fulfilling clusters had estimated angles significantly greater than $0^{\circ}$ (p-value < 0.025; Fig. 7b) and less than $45^{\circ}$ (p-value < 0.025; Fig. 7b), for which 9 normality was not rejected for (Henze-Zeckler p-value > 0.05). Fig. 7c shows the difference between original and ARIMA estimated cluster angles. 
%</SelectivelyDifferencedBoxJenkinsCorrelatedSummary>

%<*SelectivelyDifferencedBoxJenkinsTestsPassedEigenvalueSummary>
The factor analysis model from above can be applied to the 44 correlated criteria fulfilling clusters, as the factor correlation matrices of each cluster only had one eigenvalue greater than 1. 
%</SelectivelyDifferencedBoxJenkinsTestsPassedEigenvalueSummary>

>>>>

%<*ClusterCorrelationSummary0>
Spike pairs within Stage 1 cluster ellipses were linearly correlated above chance levels (Fisher's method: $\chi^{2}(7812, N = 3906) = 15150.9, $ $p < .001$; Fig. 3e). 
%</ClusterCorrelationSummary0>

%<*ClusterCorrelationSummary0p1>
Moreover, for Stage 2 clusters, spike pair differences were correlated with the spike time of the first unit in the pair (Fisher's method: $\chi^{2}(324, N = 162) = 7832.7, $ $p < .001$; Fig. 3g), showing that correlations were not explained by a model of the form $s_1 = s_0 + d + independent\_noise$ where $d$ is a fixed difference.
%</ClusterCorrelationSummary0p1>

%<*ClusterCorrelationSummary1>
97.5\% (158/162) of Stage 2 clusters were positively correlated (Fig. 3f). 
%</ClusterCorrelationSummary1>

%<*ClusterCorrelationSummary2>
80.9\% (131/162) of the Stage 2 clusters had correlations of higher significance than correlations calculated for all unclustered first spike pairs in the original response distribution (Fig. 3h). Moreover, 72.8\% (118/162) of the original response distributions from which Stage 2 clusters were extracted were not correlated significantly (p>0.005) (Fig. 3h). 
%</ClusterCorrelationSummary2>

%<*AngledClustersUniquePairsSummary>
A total of 162 unique Stage 2 clusters were extracted from 162 unique response distributions.
%</AngledClustersUniquePairsSummary>

%<*OriginalAnglesSummary>
67.9\% (110/162) of $	heta_{45}$ angles were both significantly less than $45^{\circ}$ (p<0.025) and greater than $0^{\circ}$ (<0.025). 
%</OriginalAnglesSummary>

%<*OriginalTestsPassedSummary>
Of the 162 angled clusters, 46 were determined to be stationary and sufficiently correlated by fulfilling the criteria described in the Methods \ref{methods:angle_and_factor_analysis_criteria}, such that an underlying factor could underlie the covariation.
%</OriginalTestsPassedSummary>

%<*OriginalTestsPassedAngleSummary>
Of the 46 criteria fulfilling angled clusters, 30 had estimated angles significantly different from both $0^{\circ}$ (p<0.025; Fig. 4f) and $45^{\circ}$ (p<0.025; Fig. 4f). Of the 46 criteria fulfilling angled clusters, normality was not rejected for 20 clusters (Henze-Zirkler p > 0.05), of which 14 had an estimated angle significantly different from both $0^{\circ}$ (p-value < 0.025) and $45^{\circ}$ (p-value < 0.025).
%</OriginalTestsPassedAngleSummary>

%<*OriginalTestsPassedFactorAnalysisSummary>
Factor analysis could be applied to the 46 criteria fulfilling angled clusters, to give an estimate of the factor-conditioned cluster response distributions for the stationary angled clusters.
%</OriginalTestsPassedFactorAnalysisSummary>

%<*ClusterSingleUnitKPSSStationaritySummary>
For the 162 angled clusters, 117 of the corresponding 324 ($=2*162$) single unit cluster first spike sequences were determined as KPSS non-stationary (p < 0.05; Fig. 4d).
%</ClusterSingleUnitKPSSStationaritySummary>

%<*ClusterSingleUnitTrialIndexLRSummary>
The cluster spike times of either 0, 1 or 2 of the units in a cluster pair were correlated with trial index (Fisher's method: $\chi^{2}(648, N = 324) = 2485.1, $ $p < .001$; Fig. 4b,c).
%</ClusterSingleUnitTrialIndexLRSummary>

%<*SelectivelyDifferencedSummary>
In total, 83 clusters had differencing applied to atleast one unit. 83 of these clusters fulfilled the criteria, giving a total of 83 criteria fulfilling angled clusters (including the undifferenced stationary clusters). 
%</SelectivelyDifferencedSummary>

%<*SelectivelyDifferencedCorrelatedSummary>
In total, 53 of the 83 criteria fulfilling clusters were linearly correlated following differencing (p-value < 0.05) (Fig. 6h). 23 of the correlated criteria fulfilling clusters had estimated angles significantly greater than $0^{\circ}$ (p-value < 0.025) and less than $45^{\circ}$ (p-value < 0.025). Of these, normality was not rejected for 13 clusters (Henze-Zirkler p > 0.05). Fig. 6j shows that the distribution of estimated angles for the successfully differenced clusters. Fig. 8? shows the difference between original and differenced estimated clusters?
%</SelectivelyDifferencedCorrelatedSummary>

%<*SelectivelyDifferencedTestsPassedNewSummary>
In total, 83 clusters had differencing applied to atleast one unit. 35 of these were correlated and fulfilled the criteria. 23 of these clusters had estimated angles significantly greater than $0^{\circ}$ (p-value < 0.025; Fig 6b) and less than $45^{\circ}$ (p-value < 0.025; Fig 6b), of which 12 normality was not rejected for (Henze-Zirkler p < 0.05). These angles are similar to the original angles before differencing (Fig. 6c), showing that the differencing operation does not largely alter cluster angle. Fig. 6d shows the amount of variance explained by the pairwise correlation after trend removal. 
%</SelectivelyDifferencedTestsPassedNewSummary>

%<*SelectivelyDifferencedTestsPassedEigenvalueSummary>
The factor analysis model from above can be applied to the 35 correlated criteria fulfilling differenced clusters, as the factor correlation matrices of each cluster only had one eigenvalue greater than 1. 
%</SelectivelyDifferencedTestsPassedEigenvalueSummary>

%<*SelectivelyDifferencedBoxJenkinsDifferencedSummary>
81 of the 83 differenced clusters had an ARMA model applied to atleast one unit. Almost all ARMA models fit for differenced clusters were of the MA type (Fig 7a), as expected for the negative autocorrelations introduced by differencing. 
%</SelectivelyDifferencedBoxJenkinsDifferencedSummary>

%<*SelectivelyDifferencedBoxJenkinsCorrelatedSummary>
After the described application of the Box-Jenkins method, there were 42 criteria fulfilling clusters. Normality was not rejected for 18 of these clusters (Henze-Zirkler p > 0.05). 27 of the correlated criteria fulfilling clusters had estimated angles significantly greater than $0^{\circ}$ (p-value < 0.025; Fig. 7b) and less than $45^{\circ}$ (p-value < 0.025; Fig. 7b), for which 10 normality was not rejected for (Henze-Zeckler p-value > 0.05). Fig. 7c shows the difference between original and ARIMA estimated cluster angles. 
%</SelectivelyDifferencedBoxJenkinsCorrelatedSummary>

%<*SelectivelyDifferencedBoxJenkinsTestsPassedEigenvalueSummary>
The factor analysis model from above can be applied to the 42 correlated criteria fulfilling clusters, as the factor correlation matrices of each cluster only had one eigenvalue greater than 1. 
%</SelectivelyDifferencedBoxJenkinsTestsPassedEigenvalueSummary>

### 2020-08-10 manuscript.tex
Lunch!

\subsection{Notes}
I have begun developing several techniques to address this but I think it may prove difficult to show anything strongly conclusive either way. I can give more details of this once we are agreed on the overall objectives. We may be able to address this issue, but it may turnout to require future experiments designed to get a higher cluster count per stimulus and experimental session. A bit of help attacking this problem and working on factor analysis in 3+ dimensions would also be helpful.

@@

The dataset used for this paper is reused from an earlier study, of which analyses and data collection methods have previously been reported \cite{reyes2014laminar}. In summary, extracellular recordings were made in 15 anaesthetised Wistar rats using 16 or 128 electrode silicon probes, inserted perpendicularly into the barrel cortex. Single whisker deflections were made at various frequencies (0.1-10Hz) in blocks, each containing typically 200 trials (sweeps) with an intertrial interval of 5 s (occasionally 10, 15, or 30 s). This was repeated for each of the selected target whiskers (maximum 4). Experimental sessions varied in terms of electrode placement, stimulated whiskers, stimulation frequencies and number of trials per unique stimulus. 7 experimental sessions used single shanks with 16 electrodes, whilst 8 experimental sessions used 8 neighbouring shanks (16 electrodes per shank) covering 2-4 barrel-related columns (median 4). Units were classified as excitatory or inhibitory and assigned to a cortical layers L2/3, L4, L5A and L5B/6 (see Methods \ref{methods:spike_sorting}). Spiking activity is analysed following cortical activation onset, determined as 5.2ms following stimulus onset (Supplementary Fig. 1).

@@

This section demonstrates that the responses of some neurons exhibit depressive latency adaptation, which should be controlled for when testing for correlative structure caused by a non-adaptive state. Figure 2d shows the first spike times of 5 single units in response to repeated stimulations. The L23E unit of Fig. 2d(i) demonstrates an irregular response pattern with adaptation. Whereas the L4I unit of Fig. 2d(ii) shows a short and consistent response, the L… unit of Fig. 2d(iii) reveals a sparse and variable pattern and no obvious adaptation. The L4I and L4I units of Fig. 2d(iv, v) demonstrate more complex response-distribution structure and possible non-linear non-stationarity.

@@

This section presents an example first spike pair response distribution for the purposes of understanding. Fig. 3a shows the distribution of first spike pairs in response to a 4Hz stimulation of a single whisker for two units in L5AI and L4I of the corresponding principle column (over a 50ms window following cortical activation onset). Only responses on `conjunctive trials' (trials in which both units spiked for the single stimulus) are shown.

The orange Gaussian is centered at the mean of the first spike pairs and has widths equal to 4 times the first spike time standard deviations of each unit (on conjunctive trial), illustrating the region from which 99.9\% of first spike pairs would be predicted to fall under the assumption of independence and normality (i.e. spike time jitter). This assumption fails to capture the distribution structure of this and many other spike pair distributions (Supplementary Vid. 1) and very few different stimuli to be encoded with high-discriminability.

The spike time differences have a standard deviation of X (Fig. 3b), which is also poor for discriminable encoding.

First spike pairs in Fig. 3a appear to be drawn from one of two clusters (blue and green dots) in response to the same stimulus. The green cluster appears correlated with non-unitary gradient, suggestive of an underlying factor, which covaries first spike times, such that distributions of differences would be smaller when factor-conditioned, offering a higher power for discriminable encoding.

\subsection*{Clustering algorithm summary}
A two-stage algorithm was developed to quantify the occurrence and angle of correlated clusters. 
Stage 1 extracted well isolated \textit{`flat clusters'} defined by the $4\sigma$ boundary of flat 2D Gaussian ellipses (flat solid green ellipse in Fig. 3a; see Methods \ref{methods:flat_clusters} for details).  
% The algorithm applied the unsupervised DBSCAN clustering method to first spike response distributions for a range of values of the DBSCAN EPS free parameter. Repeated clusterings for different values of the EPS free parameter were removed from analysis. For DBSCAN clusters that were well isolated and had more than 20 elements, a bounding ellipse was defined from the means and standard deviations of points in the DBSCAN cluster (flat solid green ellipse in Fig. 3a). Points in this bounding ellipse were assigned to the `flat cluster'. 
To assess the occurrence of correlated clusters without sample bias, repeated and similar flat clusters produced by non-identical DBSCAN clusterings were left in the final group of flat clusters (Methods \ref{methods:flat_clusters}). When the algorithm was verified on surrogate control datasets (see Methods \ref{methods:correlation_surrogate_control_datasets}) correlated clusters were found at chance levels (Supplementary Fig. 3).

Stage 2 of the clustering algorithm aimed to better estimate the shape of the distributions underlying correlated flat clusters (p-value < 0.005), by estimating an `angled cluster' ellipse (angled solid green ellipse in Fig. 3a; see Methods \ref{methods:angled_clusters}). Repeats, similar and non-isolated angled clusters were discarded. 
\ExecuteMetaData[100msAnalysisOutputLatex.tex]{ClusterTimeSpanSummary} \ExecuteMetaData[AnalysisOutputLatex.tex]{AngledClustersUniquePairsSummary}
X\%\todo[color=red!40]{Fill} of angled clusters contained spikes from all conjunctive trials (Fig. 3d).

\subsection*{Correlation analysis}
\ExecuteMetaData[AnalysisOutputLatex.tex]{ClusterCorrelationSummary1}
\ExecuteMetaData[AnalysisOutputLatex.tex]{ClusterCorrelationSummary2}

@@

\subsection*{Cluster angle}
\label{results:cluster_angle}

@@

An algorithm was developed to estimate a mean angle and confidence interval for each angled cluster (see Methods \ref{methods:cluster_angle_confidence_interval} for details), to test whether the angles of angled clusters were significantly different from $45^{\circ}$ (possibly indicative of factor or trial-dependent spike time differences). The mean angles, $\theta$, of positively correlated angled clusters (between $0^{\circ}$ and $90^{\circ}$) were of interest. $\theta_{45} = min(\theta, 90-\theta)$ were calculated for analysis, as only the ordering of the units in the pair determined whether $\theta$ was above or below $45^{\circ}$.

\ExecuteMetaData[AnalysisOutputLatex.tex]{OriginalAnglesSummary}
The algorithm was tested on a surrogate control dataset to validate that the detection of such differences from $45^{\circ}$ were not artifacts of the algorithm (see Methods \ref{methods:angle_surrogate_control_dataset}; Supplementary Fig. 6). The following subsections aim to determine when trial number and/or another factor underlie such non-unitary covariation.

>>>>

% \subsection{Notes}
% I have begun developing several techniques to address this but I think it may prove difficult to show anything strongly conclusive either way. I can give more details of this once we are agreed on the overall objectives. We may be able to address this issue, but it may turnout to require future experiments designed to get a higher cluster count per stimulus and experimental session. A bit of help attacking this problem and working on factor analysis in 3+ dimensions would also be helpful.

@@

The dataset used for this paper is reused from an earlier study, of which analyses and data collection methods have previously been reported \cite{reyes2014laminar}. In summary, extracellular recordings were made in 15 anaesthetised Wistar rats using 16 or 128 electrode silicon probes, inserted perpendicularly into the barrel cortex. Single whisker deflections were made at various frequencies (0.066-10Hz) in blocks, each containing typically 200 trials (sweeps) with an inter trial interval of 5s (occasionally 10, 15, or 30 s). This was repeated for each of the selected target whiskers (maximum 4). Experimental sessions varied in terms of electrode placement, stimulated whiskers, stimulation frequencies and number of trials per unique stimulus. 7 experimental sessions used single shanks with 16 electrodes, whilst 8 experimental sessions used 8 neighbouring shanks (16 electrodes per shank) covering 2-4 barrel-related columns (median 4). Units were classified as excitatory or inhibitory and assigned to a cortical layers L2/3, L4, L5A and L5B/6 (see Methods \ref{methods:spike_sorting}). Spiking activity is analysed following cortical activation onset, determined as 5.2ms following stimulus onset (Supplementary Fig. 1).

@@

This section demonstrates that the responses of some neurons in all neuron-groups exhibit latency adaptation, which should be controlled for when testing for correlative structure caused by a non-adaptive state. Figure 2d shows the first spike times of 5 single units in response to repeated stimulations. The L23E unit of Fig. 2d(i) demonstrates an irregular response pattern with adaptation. Whereas the L4I unit of Fig. 2d(ii) shows a short and consistent response, the L… unit of Fig. 2d(iii) reveals a sparse and variable pattern and no obvious adaptation. The L4I and L4I units of Fig. 2d(iv, v) demonstrate more complex response-distribution structure and possible non-linear non-stationarity.

@@

For the purpose of understanding, this section presents an example distribution of first spike pairs for a pair of units (L5AI and L4I) in the principle column in response to a 4Hz stimulation of the corresponding principle whisker (over a 50ms window following cortical activation onset). Only responses on `conjunctive trials' (trials in which both units spiked for the single stimulus) are shown. The orange Gaussian is centered at the mean of the first spike pairs and has widths equal to 4 times the first spike time standard deviations of each unit (on conjunctive trial), illustrating the region from which 99.9\% of first spike pairs would be predicted to fall under the assumption of independence and normality (i.e. spike time jitter). This assumption fails to capture the distribution structure of this and many other spike pair distributions (Supplementary Vid. 1). 
% Moreover, very few different stimuli could be encoded with high-discriminability.

% The spike time differences have a standard deviation of X (Fig. 3b), which is also poor for discriminable encoding.

Instead, first spike pairs appear to come from one of two clusters (blue and green dots). Moveover, the green cluster appears correlated with non-unitary gradient, suggestive of some underlying continuous state, which covaries the first spike times of the two units from trial-to-trial. Moreover, the apparent non-unitary gradient of correlation suggests that the distribution of differences varies with respect to the value of this underlying state, such that the state-conditioned distribution of relative first spike differences would be more compact than the general case, thus offering a higher potential capacity for stimulus encoding if downstream neurons could respond in a state-dependent manner.

% \subsection*{Clustering algorithm summary}

\subsection*{Stage 1 Clusters - Correlation analysis}

A two-stage algorithm was developed to quantify the occurrence and angle of correlated clusters. Stage 1 of the algorithm extracted well isolated clusters defined by the $4\sigma$ boundary of flat 2D Gaussian ellipses (see Methods \ref{methods:flat_clusters}; e.g. flat solid green ellipse in Fig. 3a).
\ExecuteMetaData[AnalysisOutputLatex.tex]{ClusterCorrelationSummary0}
The algorithm was verified by confirming that the algorithm extracted correlated clusters at chance levels for multiple control data sets (see Methods \ref{methods:correlation_surrogate_control_datasets}; Supplementary Fig. 3).

% The algorithm applied the unsupervised DBSCAN clustering method to first spike response distributions for a range of values of the DBSCAN EPS free parameter. Repeated clusterings for different values of the EPS free parameter were removed from analysis. For DBSCAN clusters that were well isolated and had more than 20 elements, a bounding ellipse was defined from the means and standard deviations of points in the DBSCAN cluster (flat solid green ellipse in Fig. 3a). Points in this bounding ellipse were assigned to the `flat cluster'. 
% To assess the occurrence of correlated clusters without sample bias, repeated and similar flat clusters produced by non-identical DBSCAN clusterings were left in the final group of flat clusters (Methods \ref{methods:flat_clusters}).

@@

\subsection*{Stage 2 Clusters - Angle analysis}
\label{results:cluster_angle}

To estimate the correlation angle of the correlated Stage 1 clusters, a corresponding `Stage 2' \textit{angled} $4\sigma$ ellipse was estimated for each correlated Stage 1 cluster (see Methods \ref{methods:angled_clusters}; e.g. angled solid green ellipse in Fig 3a). Stage 2 clusters which were non-isolated, repeats or contained smaller clusters were discarded. 
\ExecuteMetaData[100msAnalysisOutputLatex.tex]{ClusterTimeSpanSummary}

\ExecuteMetaData[AnalysisOutputLatex.tex]{AngledClustersUniquePairsSummary}
\ExecuteMetaData[AnalysisOutputLatex.tex]{ClusterCorrelationSummary2}

@@

% X\%\todo[color=red!40]{Fill} of angled clusters contained spikes from all conjunctive trials (Fig. 3d).

To test whether the cluster correlation angles were different from $45^{\circ}$, an algorithm was developed to estimate the angle $\theta$ of correlation for each Stage 2 cluster (see Methods \ref{methods:cluster_angle_confidence_interval}). A final $\theta_{45} = min(\theta, 90-\theta)$ angle was used for each Stage 2 cluster, as only the ordering of the units determined whether $\theta$ was above or below $45^{\circ}$.

\ExecuteMetaData[AnalysisOutputLatex.tex]{ClusterCorrelationSummary1}
Fig. 4a shows the estimated $\theta_{45}$ angles (with 95\% confidence intervals) for all positively correlated Stage 2 clusters.
\ExecuteMetaData[AnalysisOutputLatex.tex]{OriginalAnglesSummary}
This initial finding that shows that the distribution of relative first spike differences between two units depends on some some underlying processes.
Following sections will aim to determine when neural adaptation and/or non-adaptative underlying states underlie such positive non-unitary correlations. 
To validate that such differences from $45^{\circ}$ were not artifacts of the algorithm, the algorithm was tested on a surrogate control dataset (see Methods \ref{methods:angle_surrogate_control_dataset}; Supplementary Fig. 6).

The non-unitary correlations cannot be explained by a model in which the second spiking neuron spikes a fixed time difference after the first. 
\ExecuteMetaData[AnalysisOutputLatex.tex]{ClusterCorrelationSummary0p1}
This strengthens the case that the spike timing correlations between pairs of neurons are co-modulated by some underlying processes.
